---
title: "Advanced iteration examples"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 6
    
always_allow_html: true
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
# figs folder
fs::dir_create("figs")
# data folder
fs::dir_create("data")
# docs folder
fs::dir_create("docs")
# chunk options
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/", # location of files
  fig.height = 5.5, # height of figures
  fig.width = 8 # width of figures
)
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)

# nyt1 <- read_csv("data/dds_ch2_nyt/nyt1.csv")
# nyt1 <- nyt1 %>% dplyr::slice_sample(n = 100)
# readr::write_csv(x = nyt1, path = "data/dds_ch2_nyt/nyt1.csv")
# nyt2 <- read_csv("data/dds_ch2_nyt/nyt2.csv")
# nyt2 <- nyt2 %>% dplyr::slice_sample(n = 75)
# readr::write_csv(x = nyt2, path = "data/dds_ch2_nyt/nyt2.csv")
# nyt3 <- read_csv("data/dds_ch2_nyt/nyt3.csv")
# nyt3 <- nyt3 %>% dplyr::slice_sample(n = 85)
# readr::write_csv(x = nyt3, path = "data/dds_ch2_nyt/nyt3.csv")
# nyt4 <- read_csv("data/dds_ch2_nyt/nyt4.csv")
# nyt4 <- nyt4 %>% dplyr::slice_sample(n = 85)
# readr::write_csv(x = nyt4, path = "data/dds_ch2_nyt/nyt4.csv")
# nyt5 <- read_csv("data/dds_ch2_nyt/nyt5.csv")
# nyt5 <- nyt5 %>% dplyr::slice_sample(n = 50)
# readr::write_csv(x = nyt5, path = "data/dds_ch2_nyt/nyt5.csv")
# csv_files <- fs::dir_ls(path = "data/dds_ch2_nyt", glob = "*.csv")
# nyt_csv_list <- purrr::map(.x = csv_files, .f = read_csv)
# readr::write_rds(x = nyt_csv_list, path = "data/nyt_csv_list.rds")
```



## Importing multiple .csv files

Consider this example , where we have multiple .csv files in the `data` folder. All of these files are similar (same column names, but different numbers of rows) and we want them all loaded into our RStudio environment.

```{r dds_ch2_nyt}
fs::dir_tree("data")
```

Assuming we've imported a single file and discovered some wrangling that needs to be done for each file. We've packages these wrangling steps into a custom function called `clean_nyt_data`.

```{r clean_nyt, message=FALSE}
clean_nyt <- function(file) {
                    nyt <- read_csv(file)
                    nyt %>%
                    dplyr::mutate(
                        age_group = case_when( # create age_group variable
                                        Age < 18 ~ "<18",
                            Age >= 18 & Age < 25 ~ "18-24",
                            Age >= 25 & Age < 35 ~ "25-34",
                            Age >= 35 & Age < 45 ~ "35-44",
                            Age >= 45 & Age < 55 ~ "45-54",
                            Age >= 55 & Age < 65 ~ "55-64",
                            Age >= 65 ~ "65+"),
                        CTR = Clicks/Impressions, # create CTR variable
                        Female = case_when( # create new Female variable
                                Gender == 0 ~ "Male",
                                Gender == 1 ~ "Female",
                    TRUE ~ as.character(Gender)))
}
# test this on single file,
clean_nyt("data/dds_ch2_nyt/nyt1.csv") %>% head()
```

Ok, now we want to combine our cleaning and importing into a single for loop to iterate through each .csv file. This takes the following steps,

1. Build a file path to each .csv

```{r nyt_files}
# build a vector of files
nyt_files <- base::dir("data/dds_ch2_nyt")
nyt_files <- base::paste0("data/", "dds_ch2_nyt/", nyt_files)
glimpse(nyt_files)
```

2. Construct a `for loop` that iterates over each file, imports it, and adds the `id` column so we can keep track of each file origin.

```{r nyt_data-for-loop, message=FALSE, warning=FALSE}
# Build my_nyt_data with a for loop
my_nyt_data <- NULL
for (file in nyt_files) { # for every file in nyt_files...
    # wrangle with clean_nyt()
    temp <- clean_nyt(file)  
    # add an id column (but remove .csv)
    temp$id <- sub(".csv", "", file)
     # then stick together by rows
    my_nyt_data <- rbind(my_nyt_data, temp)
}
my_nyt_data %>% glimpse()
```

Now when we check the output of the `for loop`, we should have 5 different `id`s.

```{r count-ids}
unique(my_nyt_data$id) %>% length()
```

We could also do the same thing with a `list`,  and might look like this,

1. put the files in a list

```{r nyt_files_list}
files <- as.list(x = nyt_files) %>% purrr::set_names()
# get path to files and store in list
nyt_files_list <- files
# check the length
length(nyt_files_list)
```

2. Construct the `for loop` to import the files in `nyt_files_list`, with some minor changes like `seq_along()` and bracket subsetting.

```{r for-loop, message=FALSE}
# the for loop for all of the items in the files list
for (i in seq_along(files)) {
    # apply read_csv to all the items in files
  nyt_files_list[[i]] <- readr::read_csv(file = files[[i]])

}
```



The number of imported files is contained in the file names.

```{r head-nyt1.csv}
head(nyt_files_list$`data/dds_ch2_nyt/nyt1.csv`)
```

### Using map

The basic method for `purrr` map is,

1. Do it for one element  
2. Turn it into a recipe   
3. Use `map()` to do it for all elements   

So the obvious 'do it for one' here to import and wrangle the data (which we have done above with `clean_nyt`). So the recipe is,

```r
 ~ clean_nyt( .x$`data/dds_ch2_nyt/nyt1.csv`)
```

First we'll store the files in a `fs_path` from the `fs` package. These are great for keeping track of files.

```{r dir_ls-csv_files}
csv_files <- fs::dir_ls(path = "data/dds_ch2_nyt", glob = "*.csv")
str(csv_files)
```

Using `map`, this would look like so,

```{r}
csv_files <- fs::dir_ls(path = "data/dds_ch2_nyt", glob = "*.csv")
nyt_csv_list_map <- purrr::map(.x = csv_files, .f = read_csv)
```


Below is a list with a small sample of data.frames from [Doing Data Science](https://www.oreilly.com/library/view/doing-data-science/9781449363871/). The information on these data is provided below,

> Each dataset represents one (simulated) day’s worth of ads shown and clicks recorded on the New York Times home page in May 2012. Each row represents a single user. There are five columns: `Age`, `Gender` (0 = female, 1 = male), `Impressions`, `Clicks`, and `Signed_In`.

The code chunk below imports these data as `nyt_csv_list`.

```{r import-nyt_csv_list}
nyt_csv_list <- readr::read_rds(path = "data/nyt_csv_list.rds")
str(nyt_csv_list)
```

We can see each `data.frame` in the list has the same five variables.

```{r names-nyt1}
names(nyt_csv_list$nyt1)
```

Map always returns a list,
