---
title: "Iteration best practices"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 6
    
always_allow_html: true
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
library(tidyverse)
library(devtools)
# figs folder
fs::dir_create("figs")
# data folder
fs::dir_create("data")
# docs folder
fs::dir_create("docs")
# chunk options
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/", # location of files
  fig.height = 5.5, # height of figures
  fig.width = 8 # width of figures
)
# knit options
knitr::opts_knit$set(
  width = 78,
  progress = FALSE
)
# base options
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  max.print = 999999,
  scipen = 100000000
)
```

# Motivation

> `lapply`, `map`, `purrr` or good old for-loops. After trial and error I seem to be able to get my `for loops` to work, but I always wonder if there is a better way?

There is nothing wrong with `lapply` and others per se, but `purrr` allows for cleaning code (easier to read), and can make many iteration activities easier because it requires less typing and runs a bit faster.

Below are examples of using `purrr` to solve common iteration problems I've encountered. Not all of these might be useful to you, but I am hoping these provide enough information for you to have a basic understanding of `purrr` and why it's preferred to most `for loops`.

## Subsetting lists

It's important to remember how to access the elements in a list. These can be vectors, data.frames/tibbles, or other lists.

Let's start by creating a list of `scores` of that are randomly selected from the `rnorm()` distribution. Each vector has a different length, so it wouldn't make sense to put these into a `data.frame`.

```{r scores-list}
scores <- list(
  score_01 = round(rnorm(n = 8, mean = 7, sd = 2), 0),
  score_02 = round(rnorm(n = 9, mean = 4, sd = 1), 0),
  score_03 = round(rnorm(n = 10, mean = 6, sd = 2), 0),
  score_04 = round(rnorm(n = 11, mean = 3, sd = 1), 0),
  score_05 = round(rnorm(n = 10, mean = 8, sd = 2), 0),
  score_06 = round(rnorm(n = 12, mean = 9, sd = 2), 0),
  score_07 = round(rnorm(n = 7, mean = 2, sd = 2), 0),
  score_08 = round(rnorm(n = 10, mean = 9, sd = 3), 0),
  score_09 = round(rnorm(n = 5, mean = 3, sd = 2), 0),
  score_10 = round(rnorm(n = 4, mean = 1, sd = 1), 0)
)
```

If we want to get `score_05` from the `scores`, we have a few options. Single brackets return the list, dollar sign and double brackets return a vector.

```{r subsetting-lists}
str(scores[1])
str(scores$score_01)
str(scores[[1]])
```

Note that we can do this with named lists. If the list elements are unnamed, we can still use the double brackets to extract elements from a list.

## Applying functions across the elements of a list

Assume we'd like to get the median value for each score in `scores`. `purrr::map()` allows us to do this easily by simply providing the list (`.x`) and function (`.f`) arguments.


```{r map-basics}
map(.x = scores, .f = median)
```

`purrr::map()` always returns a list, but there are alternative functions, too.

### Alternatives to map()

We can use `purrr::map_dbl()` to see the results as a vector.

```{r map_dbl-basics}
map_dbl(.x = scores, .f = median)
```


To see the full list of `purrr::map()` functions, type `map_` into RStudio and they should populate using tab completion.

## Mappers

Mappers are anonymous functions with a one-sided formula.

### Types of functions

Below is a quick overview of two different types of functions.

#### Classical functions

Below we create a classical 'named' function called `my_fun()`, which gets the `sd` (standard deviation) and `round`s it.

```{r classical-functions}
my_fun <- function(x) {
  round(sd(x))
}
# apply with map_dbl
map_dbl(.x = scores, .f = my_fun)
```

#### Lambda functions

And these are lambda 'unnamed' functions, where the function is created 'on the fly', and included inside the `map_` function.

```{r lambda-function}
map_dbl(.x = scores, .f = function(x) {
  round(sd(x))
  })
```

#### Mapping single parameters

Mappers use the .x to refer to the parameter of the function. You can also use `.` or `..` to refer to `.x`.

```{r mapping-to-parameters}
map_dbl(scores, ~ round(sd(.x)))
map_dbl(scores, ~ round(sd(.)))
map_dbl(scores, ~ round(sd(..1)))
```

I suggest you stick with `.x` because its easy to get lost.

#### Mapping multiple parameters

If you want to refer to more than one parameter, you can refer to the second parameter as `.y`. We'll create another list called `more_scores`.

```{r more_scores-list}
more_scores <- list(
  score_11 = round(rnorm(n = 8, mean = 10, sd = 3), 0),
  score_12 = round(rnorm(n = 9, mean = 11, sd = 3), 0),
  score_13 = round(rnorm(n = 10, mean = 12, sd = 5), 0),
  score_14 = round(rnorm(n = 11, mean = 11, sd = 1), 0),
  score_15 = round(rnorm(n = 10, mean = 11, sd = 2), 0),
  score_16 = round(rnorm(n = 12, mean = 10, sd = 1), 0),
  score_17 = round(rnorm(n = 7, mean = 15, sd = 3), 0),
  score_18 = round(rnorm(n = 10, mean = 11, sd = 2), 0),
  score_19 = round(rnorm(n = 5, mean = 15, sd = 3), 0),
  score_20 = round(rnorm(n = 4, mean = 13, sd = 4), 0)
)
more_scores
```

Below we combine `scores` and `more_scores` with `.x` and `.y`

```{r two-parameters}
map2(scores, more_scores, ~ .x + .y)
```

If there are more than two parameters, we can use `pmap` and combine the lists with `list`

```{r even_more_scores-list}
even_more_scores <- list(
  score_21 = round(rnorm(n = 8, mean = 21, sd = 10), 0),
  score_22 = round(rnorm(n = 9, mean = 20, sd = 12), 0),
  score_23 = round(rnorm(n = 10, mean = 23, sd = 11), 0),
  score_24 = round(rnorm(n = 11, mean = 22, sd = 13), 0),
  score_25 = round(rnorm(n = 10, mean = 21, sd = 10), 0),
  score_26 = round(rnorm(n = 12, mean = 25, sd = 12), 0),
  score_27 = round(rnorm(n = 7, mean = 22, sd = 14), 0),
  score_28 = round(rnorm(n = 10, mean = 20, sd = 12), 0),
  score_29 = round(rnorm(n = 5, mean = 24, sd = 15), 0),
  score_30 = round(rnorm(n = 4, mean = 20, sd = 10), 0)
)
even_more_scores
```

```{r pmap}
pmap(.l = list(scores, more_scores, even_more_scores), ~ ..1 + ..2 + ..3)
```

### Creating mapper functions

To create a mapper function, we can use `purrr::as_mapper()` function.

```{r my_fun_mapper}
my_fun_mapper <- as_mapper(~ round(sd(.x)))
```

These are easier to use, more concise, and should be used whenever you catch yourself copying and pasting functions.

```{r use-my-fun-mapper}
map_dbl(scores, my_fun_mapper)
```

## Setting names

`purrr` has a great function for setting names called `set_names()`. It works by attaching names to the elements of a list.  

```{r weekday_abbr}
weekday_abbr <- c("Mon", "Tues", "Wed", "Thurs", "Fri", "Sat", "Sun")
days <- c(1, 2, 3, 4, 5, 6, 7)
purrr::set_names(days, weekday_abbr)
```

If we have multiple elements, we can combine them in a `list` and apply the names.

```{r days2-days3}
days2 <- c(1, 2, 3, 4, 5, 6, 7)
days3 <- c(1, 2, 3, 4, 5, 6, 7)
all_days <- list(days, days2, days3)
map(all_days, ~set_names(.x, weekday_abbr))
```

### The `keep()` function

`purrr::keep()` retains elements that satisfy conditions in a list. What if we want to keep only the `scores` that are less than `20`?

```{r keep}
purrr::keep(scores, ~ sum(.x) < 20)
```

We can also create a `mapper` with the condition and apply it to `scores`.

```{r cut_off-mapper}
cut_off <- as_mapper(~ sum(.x) < 20)
keep(scores, cut_off)
```

### The `discard` function

The counterpart to `keep` is `discard`, and it's used to drop the elements of a list that satisfy a condition. Below we use it to remove the `scores` who aren't less than `20`.

```{r discard}
discard(.x = scores, .p = cut_off)
```

Below is a list of all-time home run hitters in the MLB according to [baseball-reference.com](https://www.baseball-reference.com/leaders/HR_career.shtml).

```{r sluggers}
sluggers <- list(rank = factor(x = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                               labels = c("first", "second", "third", "fourth",
                               "fifth", "sixth", "seventh", "eighth", "ninth",
                               "tenth"), ordered = TRUE),
                 player = c("Barry Bonds", "Hank Aaron", "Babe Ruth",
                            "Alex Rodriguez", "Willie Mays", "Albert Pujols",  
                            "Ken Griffey Jr.", "Jim Thome", "Sammy Sosa",
                            "Frank Robinson"),
                 home_runs = c(762L, 755L, 714L, 696L, 660L, 659L, 630L, 612L,  
                               609L, 586L),
                 bats = c("L", "R", "L", "R", "R", "R", "L", "L", "R", "R"))
str(sluggers)
```

### Predicate functions with `keep` and `discard`

If we wanted to keep the vectors in `sluggers` are `factors`, we can use `keep` and `is.factor`

```{r keep-predicates}
keep(sluggers, is.factor)
```

If we wanted to discard all the numeric vectors in `sluggers`, we can use `discard` and `is.numeric`.

```{r discard-predicates}
discard(sluggers, is.numeric)
```

### Predicate functions with `every()` and `some()`

If we wanted to check and if all of the elements in `sluggers` are character variables, we can use `every()`

```{r every}
every(.x = sluggers, is.character)
```

And if we wanted to check and see if at least one element in `sluggers` is numeric, we can combine `is.numeric` with `some()`

```{r some}
some(.x = sluggers, is.numeric)
```

These come in handy when we have larger lists with more objects and we're evaluating what they contain.

## More complicated lists

Next we'll be using some twitter data from the `fivethirtyeight` and `fivethirtyeightdata` packages. These contain the data used for articles from [FiveThirtyEight](https://fivethirtyeight.com/). 

The code chunk below load the two twitter datasets and combines them in a list `twitter_538`

```{r twitter_538}
trump_twitter <- fivethirtyeight::trump_twitter
twitter_presidents <- fivethirtyeightdata::twitter_presidents
twitter_538 <- list(trump_twitter = trump_twitter,
                    twitter_presidents = twitter_presidents)
```

### Viewing lists with `listviewer`

Lists are a common object when using `purrr` because they can contain multiple object types. The `jsonedit` function from `listviewer` is great for viewing lists,

```{r jsonedit}
library(listviewer)
listviewer::jsonedit(twitter_538)
```

These two datasets contain various forms of twitter data:

- `trump_twitter`: A data frame with 448 rows representing tweets and 3 variables from ["The World's Favorite Donald Trump Tweets"](https://fivethirtyeight.com/features/the-worlds-favorite-donald-trump-tweets/)

- `twitter_presidents`:  A data frame with 6439 rows describing individual tweets and 8 variables from ["The Worst Tweeter in Politics Isn't Trump"](https://fivethirtyeight.com/features/the-worst-tweeter-in-politics-isnt-trump/)

We're going to use these to explore other iteration problems that `purrr` can handle.

### Evaluating `twitter_538` contents with `every()` and `some()`

Just like above, we'll use the `some()` to see if any of the elements in the `data.frame`s are dates with `is.POSIXct`. 

```{r map-keep-is.character}
map(twitter_538, ~ some(.x, lubridate::is.POSIXct))
```

We can see both `some()` is checked against both datasets in `twitter_538`. 

### Twitter MLB data

We'll collect some Twitter baseball data from the `rtweet` package to demonstrate how this might work with a data frame you've collected from Twitter. The code chunk below collects data from the `"NHL"` hashtag, 

```{r import-sports_tweets-param, echo=FALSE}
# this is here in case you've used up your API key and the code chunk below 
# won't run
sports_tweets <- readr::read_rds("data/2020-08-17/sports_tweets.rds")
nhl_tweets <- sports_tweets$nhl_tweets
mlb_tweets <- sports_tweets$mlb_tweets
```

```{r rtweet-mlb-data, eval=FALSE}
library(rtweet)
## search for 18000 tweets using the mlb hashtag
nhl_tweets <- rtweet::search_tweets(q = "NHL", n = 1000, include_rts = FALSE)
mlb_tweets <- rtweet::search_tweets(q = "MLB", n = 1000, include_rts = FALSE)
```

We'll put both data sets in a list called, `sports_tweets`.

```{r sports_tweets}
sports_tweets <- list(nhl_tweets, mlb_tweets) %>% 
  set_names(nm = c("nhl_tweets", "mlb_tweets"))
```

Let's assume we want to save this in a time-stamped folder in case we need to use it later. Let's store these items in a list, and then go about exporting them as `.rds` files.

```{r export-sports_tweets}
# create a time-stamped folder path
fs::dir_create(paste0("data/", base::noquote(lubridate::today()), "/"))

output_rds <- function(files, names){ 
  # create identical folder path
    folder_path <- paste0("data/", base::noquote(lubridate::today()), "/")
    # export each item in list with folder path + names + extension
    readr::write_rds(data, paste0(folder_path, names, ".rds"))
}

purrr::quietly(purrr::pmap(.l = list(files = sports_tweets,
                      names = names(sports_tweets)), .f = output_rds))
```

Now we'll check the contents of the data folder, we'll see these have been exported into separate data files. 

```{r check-twitter-export}
fs::dir_tree("data")
```

## Dealing with list columns

Now let's keep the columns in `mlb_tweets` that are lists and look at it with `listviewer`

```{r mlb_list_cols}
mlb_list_cols <- keep(mlb_tweets, is.list)
listviewer::jsonedit(mlb_list_cols)
```

It looks like 17 of these columns are lists, but these are hard to view and manipulate in this format. 

## Working with list columns

This section is directly from [Jenny Bryan's `purrr` tutorial](https://jennybc.github.io/purrr-tutorial/index.html), 

> To work comfortably with list-columns, you need to develop techniques to:  
> **Inspect.** What have I created?   
> **Index.** How do I pull out specific bits by name or position?   
> **Compute.** How do I operate on my list-column to make another vector or list-column?   
> **Simplify.** How do I get rid of this list-column and back to a normal data frame?

Let's assume we're only interested in tweets from the `Twitter for Android` and `Twitter for iPhone` sources, and only interested in tweets related to Oakland gambling. 

We'll use two functions from `stringr` to get some tweets from the two sources that have text of interest to us. 

```{r oakland_tweets}
# define regular expression patterns
tweets_regexp <- "Athletics|Oakland|#GamblingTwitter"
source_regexp <- "Twitter for Android|Twitter for iPhone"

oakland_tweets <- mlb_tweets %>%
  # create new versions of source and text (twords)
  mutate(source = str_extract(source, source_regexp),
         twords = str_extract_all(text, tweets_regexp)) %>% 
  # grab only four cols
  dplyr::select(source, 
                text,
                created_at, 
                twords) %>% 
  # remove the missing sources
  filter(!is.na(source))

listviewer::jsonedit(oakland_tweets)
```

Ok, now we have a smaller `data.frame` of tweets about the Oakland A's. Next we'll add a variable for the hour, and the `start` of each word.

```{r n-created_at-start}
oakland_tweets <- oakland_tweets %>%
  mutate(created_at = lubridate::as_datetime(created_at),
         hour = lubridate::hour(created_at),
         start = gregexpr(tweets_regexp, twords))
```

Let’s isolate tweets created after 1pm in which there’s an `tword` that starts within the first 10 characters.

```{r isolate-tweets}
oakland_tweets %>%
  filter(hour > 13, # after 1:00pm
         between(map_int(start, min), 0, 10)) # within the first 10 chars
```

We will add to this filter and find the tweets that contain `Oakland` with `map_lgl()` and `all`.

```{r}
oakland_tweets %>%
  filter(hour > 13, # after 1:00pm
         between(map_int(start, min), 0, 10), # within the first 10 chars
         map_lgl(twords, ~ all(c("Oakland") %in% .x))) %>%  
  unnest_auto(twords) %>% 
  unnest_auto(start)
```

We can see a new row has been create for each new `tword`. 

